Hospital Information Technology and Physician Response

### 2021-09-17 ReadingGroupPres.tex
Updates from Overleaf

\begin{frame}{The Beginning of HIT}
\centering
    \includegraphics[scale=.6]{graphics/News Clip4.PNG}
\end{frame}

\begin{frame}[fragile]{HIT has had Great (Expected) Potential in Healthcare}
\begin{alertblock}{Cost Saving}
\begin{itemize}
    \item A 2005 estimate finds possible cost reduction of hundreds of billions of dollars (\cite{hillestad2005})
\end{itemize}
\end{alertblock}

@@

\begin{frame}[fragile]{What does this mean for physicians?}
\begin{itemize}
    \item Day-to-day, this means physicians spend more time entering information into a computer and learning a new software
    \item Loss of autonomy as EHRs have progressed
    \item Physicians have experienced significant burn-out from EHRs
\end{itemize}

@@

\begin{frame}[noframenumbering]{Physician Burnout}
\begin{center}
    \includegraphics[scale=.3]{graphics/News Clip3.PNG}
\end{center}
\end{frame}

@@

\begin{frame}{Branches of Literature}
\centering
\includegraphics[scale=.45]{graphics/Contribution_litgraphic.PNG}

@@

Measures of physician labor market behavior
\begin{itemize}
    \item Number of shared patients with a hospital: CMS PSPD (2009-2015)
    \item Total services: Medicare Provider Utilization and Payment Data (starts in 2012)
\end{itemize}
\end{frame}

@@

\begin{itemize}
    \item Main variable (binary for whether a hospital uses an EHR) comes from AHA survey
    \item AHA IT Supplement also has information on EHRs that can help with documentation or decision making
    \item Meaningful use data gives information on which hospitals received a subsidy for achieving "meaningful use"
\end{itemize}
\end{frame}

@@

Physician level characteristics
\begin{itemize}
    \item Medical school graduation date, gender: Physician Compare
\end{itemize}
\end{frame}

@@

\begin{frame}{Summary Statistics}
\centering
\includegraphics[scale=.45]{Objects/sumstats_physician_table.pdf}\\
\vspace{5mm}
\includegraphics[scale=.45]{Objects/sumstats_pair_table.pdf}\\
\vspace{3mm}
\includegraphics[scale=.45]{Objects/sumstats_hospital_table.pdf}
\end{frame}

\begin{frame}{Variation in EHR Adoption}
\centering
\includegraphics[scale=.55]{Objects/TYP_plot_hospEHR_year.pdf}
\end{frame}

@@

\begin{frame}{Original Plan: Difference in Difference}
\begin{itemize}
    \item Treatment: physicians who are exposed to an EHR at one of their hospitals in some year
    \item Control: physicians who are not exposed to a hospital at \textbf{any} hospital in any year
    \begin{itemize}
        \item  Issues with control group because of spillovers
    \end{itemize}
    \item Could still try to attempt this by imposing a control group where the physician has not \textbf{yet} been exposed to an EHR at any hospital
\end{itemize}
\end{frame}

\begin{frame}{Other Ideas}
\begin{itemize}
    \item Aggregate to physician level dataset and do analysis more along the lines of the graph I showed you earlier
    \item Have a continuous measure of EHR use that captures share of exposure to EHRs
\end{itemize}
\end{frame}

\begin{frame}{Miscellaneous Points}
\begin{itemize}
    \item Want to dig deeper into meaningful use measures
    \item Try to get location measurements to think about where physicians are working
    \item Bring in other measures of work from shared patient data
\end{itemize}

@@

\begin{frame}{Variation in Meaningful Use}
    \centering
    \includegraphics[scale=.5]{Objects/TYP_plot_hospmeanuse_year.pdf}
\end{frame}

>>>>

\begin{frame}{The Potential of HIT}
\centering
    \includegraphics[scale=.6]{graphics/News Clip4.PNG}
\end{frame}

\begin{frame}[fragile]{HIT had Great (Expected) Potential in Healthcare}
\begin{alertblock}{Cost Saving}
\begin{itemize}
    \item A 2005 estimate finds possible cost reduction of hundreds of billions of dollars (\cite{hillestad2005})
\end{itemize}
\end{alertblock}

@@

\begin{frame}[fragile]{What does this mean for physicians?}
\begin{itemize}
    \item Initially, this means physicians spend more time entering information into a computer and learning a new software
    \item Loss of autonomy as EHRs have progressed
    \item Physicians have experienced significant burn-out from EHRs
\end{itemize}

@@

\begin{frame}[noframenumbering]{Physician Burnout}
\begin{center}
    \includegraphics[scale=.45]{graphics/News Clip3.PNG}
\end{center}
\end{frame}

@@

\begin{frame}{Branches of Literature}
\centering
\includegraphics[scale=.5]{graphics/Contribution_litgraphic.PNG}

@@

Measures of physician labor market behavior
\begin{itemize}
    \item Number/percent of shared patients physician $i$ has with hospital $h$: \underline{CMS Shared Patient Data} (2009-2015)
    \item Total Part B claims for physician $i$: \underline{Medicare Provider Utilization and Payment Data} (starts in 2012)
\end{itemize}
\end{frame}

@@

\begin{itemize}
    \item Main variable (binary for whether a hospital uses an EHR) comes from AHA survey
    \item AHA IT Supplement also has information on EHRs that can help with documentation or decision making
    \item (\textit{Considering Later}) Meaningful use data gives information on which hospitals received a subsidy for achieving "meaningful use"
\end{itemize}
\end{frame}

@@

Physician level characteristics
\begin{itemize}
    \item Medical school graduation date, gender: \underline{Physician Compare}
\end{itemize}
\end{frame}

@@

\begin{frame}{Summary Statistics}
\centering

\includegraphics[scale=.8]{Objects/sumstats_pair_table.pdf}

\includegraphics[scale=.8]{Objects/sumstats_physician_table.pdf}

\begin{frame}{Summary Statistics}
\centering
\includegraphics[scale=.8]{Objects/sumstats_hospital_table.pdf}

\begin{frame}{Variation in EHR Adoption}
\centering
\includegraphics[scale=.53]{Objects/TYP_plot_hospEHR_year.pdf}
\end{frame}

\begin{frame}{Looking at Shares when EHRs are Adopted}
    \centering
    \includegraphics[scale=.5]{Objects/relyear_treatment.pdf}
\end{frame}

@@

\begin{frame}{Initial Plan: Difference in Difference}
\begin{itemize}
    \item Interested in two treatment effects
    \begin{enumerate}
        \item The effect of EHR at hospital $h$ on share at hospital $h$
        \item The effect of EHR at hospital $h$ on share at hospital $-h$
    \end{enumerate}

    \vspace{2mm}

    \item My idea is to estimate these two separately, but there are issues because of spillover effects (if hospital $-h$ implements an EHR)

    \vspace{2mm}

    \item Control group will be physicians who have not yet been exposed to an EHR at \textbf{any} hospital at time $t$
\end{itemize}
\end{frame}

\begin{frame}{Other Ideas}
\begin{itemize}
    \item Aggregate to physician level dataset if there is insufficient data for the control group
    \item Have a continuous measure of EHR use that captures share of exposure to EHRs
    \item IV: bandwidth, state privacy laws
\end{itemize}
\end{frame}

\begin{frame}{Other Points}
\begin{itemize}
    \item Want to dig deeper into how a hospital qualified for meaningful use subsidies
    \item Physician location
    \item Bring in other measures of work from shared patient data
\end{itemize}

@@
